Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-25 fully-abstract-comp/technical-report.tex
Updated call and xjmp semantics to be more 'relaxed'
updated call with a condition related to the seal

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
\newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}

@@

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\begin{align*}
  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
                                     \arraycolsep=0pt
                                     \begin{array}[t]{rl}
                                       \Phi & \updReg{r_1}{w_1}\\
                                            & \updReg{r_2}{w_2}\\
                                            & \updReg{\pcreg}{c_\code}\\
                                            & \updReg{\rdata}{c_\data}
                                     \end{array} &
                                     \begin{array}[t]{l}
                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
                                       \nonExec{c_\data} \tand \\
                                       w_1 = \linCons{\Phi(r_1)} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}
                                     \end{array}
                                     \\&\\
                                     \sourcecolor\left.
                                       \arraycolsep=0pt
                                       \begin{array}[t]{rl}
                                         \Phi' & \updReg{r_2}{0} \\
                                               & \updReg{\pcreg}{\opc}\\
                                               & \updReg{\rdata}{0}\\
                                               & \updReg{\rstk}{c_\stk}
                                       \end{array}\right.& 
                                     \sourcecolor\left.
                                       \begin{array}[t]{l}
                                         \Phi(r_1) = \retptrc(\sigma) \tand \\
                                         \Phi(r_2) = \retptrd(\sigma) \tand \\
                                         \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_} \tand \\
                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
                                         c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
                                         \dom(\ms_{\stk,\priv} \uplus \ms_\stk) = [\baddr_{\stk,\priv},\eaddr_\stk] \tand \\
                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
                                       \end{array}\right.
                                     \\
                                     \failed & \totherwise
                                   \end{cases}
\end{align*}

@@

\noindent\textbf{call}\\
\lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
\lau{19-09-2017: In which cases do we have undefined here?}
\begin{align*}
  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & 
                                               \begin{cases}
                                                 \Phi'
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{rl}
                                                   &\updReg{r_1}{w_1} \\
                                                        &\updReg{r_2}{w_2} \\
                                                        &\updReg{\pcreg}{c_1}\\
                                                        &\updReg{\rdata}{c_2}\\
                                                        &\updReg{\rstk}{c_\stk}\\
                                                        &\updReg{r_{\var{ret,c}}}{\retptrc(\sigma_?)}\\
                                                        &\updReg{r_{\var{ret,d}}}{\retptrd(\sigma_?)}
                                                 \end{array}
                                                 & 
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{l}
                                                   \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
                                                   \Phi(r_2) = \sealed{c_2,\sigma_1} \tand \\
                                                   \sigma_1 = \sigma_2 \tand \\
                                                   \nonExec{c_2} \tand\\
                                                   \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
                                                   \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
                                                   \baddr_\stk \leq \aaddr_\stk \leq \eaddr_\stk \tand \\
                                                   \dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk] \tand\\
                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\baddr_\stk,\aaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\aaddr_\stk + 1,\eaddr_\stk]} \tand\\
                                                   \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                                   \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
                                                   c_\stk = \stkptr{\aaddr_\stk+1,\eaddr_\stk,\aaddr_\stk+1} \tand \\
                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk)\tand\\
                                                   \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
                                                   w_1 = \linCons{\Phi(r_1)} \tand \\
                                                   w_2 = \linCons{\Phi(r_2)}
                                                 \end{array}
                                               \end{cases}
\end{align*}

>>>>

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
\newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\sealAss}[1]{\plainfun{sealAssignment}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}

@@

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\begin{align*}
  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
                                     \arraycolsep=0pt
                                     \begin{array}[t]{rl}
                                       \Phi & \updReg{r_1}{w_1}\\
                                            & \updReg{r_2}{w_2}\\
                                            & \updReg{\pcreg}{c_\code}\\
                                            & \updReg{\rdata}{c_\data}
                                     \end{array} &
                                     \begin{array}[t]{l}
                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
                                       \nonExec{c_\data} \tand \\
                                       w_1 = \linCons{\Phi(r_1)} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}
                                     \end{array}
                                     \\&\\
                                     \sourcecolor\left.
                                       \arraycolsep=0pt
                                       \begin{array}[t]{rl}
                                         \Phi' & \updReg{r_2}{0} \\
                                               & \updReg{\pcreg}{\opc}\\
                                               & \updReg{\rdata}{0}\\
                                               & \updReg{\rstk}{c_\stk}
                                       \end{array}\right.& 
                                     \sourcecolor\left.
                                       \begin{array}[t]{l}
                                         \Phi(r_1) = \retptrc(\sigma) \tand \\
                                         \Phi(r_2) = \retptrd(\sigma) \tand \\
                                         \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_} \tand \\
                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
                                         \ms_{\stk,\priv}(\baddr_\stk-1) = \eaddr_\stk \tand \\
                                         c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
                                       \end{array}\right.
                                     \\
                                     \failed & \totherwise
                                   \end{cases}
\end{align*}

@@

\noindent\textbf{call}\\
\lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
\lau{19-09-2017: In which cases do we have undefined here?}
This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack. We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}
\begin{align*}
  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & 
                                               \begin{cases}
                                                 \Phi'
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{rl}
                                                   &\updReg{r_1}{w_1} \\
                                                        &\updReg{r_2}{w_2} \\
                                                        &\updReg{\pcreg}{c_1}\\
                                                        &\updReg{\rdata}{c_2}\\
                                                        &\updReg{\rstk}{c_\stk}\\
                                                        &\updReg{r_{\var{ret,c}}}{\retptrc(\sigma_?)}\\
                                                        &\updReg{r_{\var{ret,d}}}{\retptrd(\sigma_?)}
                                                 \end{array}
                                                 & 
                                                 \arraycolsep=0pt
                                                 \begin{array}[t]{l}
                                                   \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
                                                   \Phi(r_2) = \sealed{c_2,\sigma_2} \tand \\
                                                   \sigma_1 = \sigma_2 \tand \\
                                                   \nonExec{c_2} \tand\\
                                                   \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
                                                   \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
                                                   \baddr_\stk \leq \aaddr_\stk < \eaddr_\stk \tand \\
                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\baddr_\stk,\aaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\aaddr_\stk + 1,\eaddr_\stk]} \tand\\
                                                   c_\stk = \stkptr{\aaddr_\stk+1,\eaddr_\stk,\aaddr_\stk+1} \tand \\
                                                   \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                                   \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk) :: \stk \tand\\
                                                   \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
                                                   \mem(\baddr) = \seal{\_,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
                                                   \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand
                                                   \sigma_? = \sealAss{a} \tand\\
                                                   w_1 = \linCons{\Phi(r_1)} \tand \\
                                                   w_2 = \linCons{\Phi(r_2)}
                                                 \end{array}
                                               \end{cases}
\end{align*}
